A Horrocks-type theorem for even orthogonal $K_2$

### 2020-03-19 dl-paper.tex
Fixes sections 2 (part 2)-3

Recall that the category of (commutative) pairs $\catname{Pairs}$ is defined as follows. Its objects are pairs $(R, I)$, in which $R$ is a commutative ring with a unit and $I$ is an ideal of $R$. A morphism of pairs $f \colon (R, I) \to (R', I')$ is, by definition, a unit-preserving ring map $f \colon R \to R'$ such that $f(I) \subseteq I'$. Notice that the mapping $(R, I) \mapsto (R \to R/I)$ defines a functor from $\catname{Pairs}$ to the morphism category $\catname{CRings}^\rightarrow$.
If $(R, I)$ is such that $R$ is a local ring with maximal ideal $I$, we call such pair a {\it local pair}.

@@

Recall that a {\it crossed module} is a triple $(M, N, \mu)$ consisting of a group $N$ acting on itself by left conjugation, an $N$-group $M$ and a map
 $\mu \colon M\to N$ of $N$-groups satisfying {\it Peiffer identity} $\mu(m) \cdot m' = m m' m^{-1}$. It can be shown that the image of $\mu$ is always
  a normal subgroup of $N$ and that the kernel of $\mu$, which we denote by $L$, is always contained in the center of $M$.

@@

It is easy to see that functors $G_n$, $P_n$ and $S_n$ satisfy \ref{req:left-exact} and \ref{req:coeq}.
By~\cite{Su82} the requirement~\ref{req:subfunc} is also satisfied for $n \geq 3$.
Finally, from~\cite[Corollary~5.4]{St71} and~\cite[Theorem~1]{LS17} it follows that $S_n$ and $P_n$ satisfy~\ref{req:uce} for $n \geq 5$.
We use the notation $\KO_i(2n, R)$ as a shorthand for $\K_i^{G_n, P_n}(R)$.

Notice that $\KO_2(2n, R) = \Ker(\St(\rD_\ell, R) \to \mathrm{SO}(2n, R))$ contains $\K_2(\rD_\ell, R)$, but the converse is not generally true. On the other hand, from~\cref{lem:lowerKgroups} it follows that $\KO_3(2n, R)$ is isomorphic to the group $K_3^{G, P}$ in which $G = G(\rD_n, -) = \mathrm{Spin}(2n, -)$ and $P = \mathrm{Epin}(2n, -)$ is its elementary subfunctor.

@@

\section{An injectivity theorem for Steinberg groups} \label{firstPart}
We start this section by recalling basic notation and facts pertaining to the theory of Grothendieck--Witt groups. Recall that this theory, developed by M.~Schlichting, is a modern broad generalization of the classical hermitian K-theory of rings. We refer the reader to~\cite[\S~2]{FRS12} and~\cite[\S~2]{AF17} for an introduction to Grothendieck--Witt groups.

@@

For the rest of this section let us fix the following notation.
Let $A$ be arbitrary commutative local ring with maximal ideal $M$ and residue field $k$.
Denote by $B = B_{A, M}$ the subring $A[X^{-1}] + M[X]$ of the ring $R = A[X, X^{-1}]$ and
by $I$ the ideal $M[X, X^{-1}]$ of $B$ (it is clear that $I$ is also an ideal of $R$).

@@

\begin{externaltheorem} \label{thm41} Suppose that $2 \in A^\times$. Then for $\ell \geq 7$ the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective. \end{externaltheorem}
\begin{proof}
 Consider the following commutative diagram with exact rows, in which the lifting $t$ is obtained from~\cref{lem:lemma32}:
\begin{equation*} \xymatrix{
 C(\rD_\ell, B, I) \ar[r]^{\lambda_B} \ar@{->>}[d]_{f} & \St(\rD_\ell, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\rD_\ell, B) \ar[r]^{\nu_B} \ar[d]_{h} & \St(\rD_\ell, k[X]) \ar[d]_{i} \\
 C(\rD_\ell, R, I) \ar[r]^{\lambda_R}         & \St(\rD_\ell, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\rD_\ell, R) \ar[r]^-{\nu_R}        & \St(\rD_\ell, k[X, X^{-1}]).
}\end{equation*}
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$
 also lies in $\Ker(\nu_B)$ and hence comes from some $b \in \St(\rD_\ell, B, I)$ via $\mu_B$.
Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\rD_\ell, R, I)$ such that $\lambda_R(c) = g(b)$. 
By~\cref{lem:prop41}, $f$ is surjective, therefore $c = f(d)$ for some $d \in C(\rD_\ell, R, I)$.
The required assertion now follows from the following calculation:
 \[ 1 = \mu_B\lambda_B(d) = tg\lambda_B(d) = t\lambda_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]
\end{proof}

>>>>

Recall that the category of (commutative) pairs $\catname{Pairs}$ is defined as follows. Its objects are pairs $(R, I)$, in which $R$ is a commutative ring with $1$ and $I$ is a (not neccessarily proper) ideal of $R$. A morphism of pairs $f \colon (R, I) \to (R', I')$ is, by definition, a unit-preserving ring map $f \colon R \to R'$ such that $f(I) \subseteq I'$. Notice that the mapping $(R, I) \mapsto (R \to R/I)$ defines a functor from $\catname{Pairs}$ to the morphism category $\catname{CRings}^\rightarrow$.
If $(R, I)$ is such that $R$ is a local ring with maximal ideal $I$, we call such pair a {\it local pair}.

@@

Recall that a {\it crossed module} is a triple $(M, N, \mu)$ consisting of the following data:
\begin{enumerate}
 \item a group $N$ acting on itself by left conjugation (i.\,e. ${}^{n}\!n' = n n' n^{-1}$).
 \item a group $M$ with a left action of $N$ (we call such a group an $N$-group and use the notation ${}^n m$ to denote the image of an element $m\in M$ under the action of $n \in N$);
 \item a homorphism $\mu \colon M\to N$ preserving the action of $N$-groups and satisfying {\it Peiffer identity} ${}^{\mu(m)}\!m' = m m' m^{-1}$.
\end{enumerate}
It can be shown that the image of $\mu$ is always a normal subgroup of $N$ and that the kernel of $\mu$, which we denote by $L$, is always contained in the center of $M$.

@@

It is easy to see that the functors $G_n$, $P_n$ and $S_n$ satisfy \ref{req:left-exact} and \ref{req:coeq}.
By~\cite{Su82} the requirement~\ref{req:subfunc} is also satisfied for $n \geq 3$.
Finally, from~\cite[Corollary~5.4]{St71} and~\cite[Theorem~1]{LS17} it follows that $S_n$ and $P_n$ satisfy~\ref{req:uce} for $n \geq 5$.
We use the notation $\KO_i(2n, R)$ as a shorthand for $\K_i^{G_n, P_n}(R)$.

Notice that $\KO_2(2n, R) = \Ker(\St(\rD_n, R) \to \mathrm{SO}(2n, R))$ contains $\K_2(\rD_n, R)$, but the converse is not generally true. On the other hand, from~\cref{lem:lowerKgroups} it follows that $\KO_3(2n, R)$ is isomorphic to the group $K_3^{G, P}$ in which $G = G(\rD_n, -) = \mathrm{Spin}(2n, -)$ and $P = \mathrm{Epin}(2n, -)$ is its elementary subfunctor.

@@

\section{An injectivity theorem for Steinberg groups} \label{firstPart}
We start this section by recalling basic notation and facts pertaining to the theory of higher Grothendieck--Witt groups. Recall that this theory, developed by M.~Schlichting, is a modern broad generalization of the classical hermitian K-theory of rings. We refer the reader to~\cite[\S~2]{FRS12} and~\cite[\S~2]{AF17} for an introduction to Grothendieck--Witt groups.

@@

For the rest of this section let us fix the following notation.
Let $A$ be an arbitrary commutative local ring with maximal ideal $M$ and residue field $k$.
Denote by $B = B_{A, M}$ the subring $A[X^{-1}] + M[X]$ of the ring $R = A[X, X^{-1}]$ and
by $I$ the ideal $M[X, X^{-1}]$ of $B$ (it is clear that $I$ is also an ideal of $R$).

@@

\begin{externaltheorem} \label{thm41} Suppose that $2 \in A^\times$. Then for $\ell \geq 7$ the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective. \end{externaltheorem}
\begin{proof}
 Consider the following commutative diagram with exact rows obtained from~\eqref{LodayRelativization}:
\begin{equation*} \xymatrix{
 C(\rD_\ell, B, I) \ar[r]^{\lambda_B} \ar@{->>}[d]_{f} & \St(\rD_\ell, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\rD_\ell, B) \ar[r]^{\nu_B} \ar[d]_{h} & \St(\rD_\ell, k[X]) \ar[d]_{i} \\
 C(\rD_\ell, R, I) \ar[r]^{\lambda_R}         & \St(\rD_\ell, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\rD_\ell, R) \ar[r]^-{\nu_R}        & \St(\rD_\ell, k[X, X^{-1}]).
}\end{equation*}
The lifting $t$ in the above diagram is obtained from~\cref{lem:lemma32}.
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$
 also lies in $\Ker(\nu_B)$ and hence comes from some $b \in \St(\rD_\ell, B, I)$ via $\mu_B$.
Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\rD_\ell, R, I)$ such that $\lambda_R(c) = g(b)$. 
By~\cref{lem:prop41}, $f$ is surjective, therefore $c = f(d)$ for some $d \in C(\rD_\ell, R, I)$.
The required assertion now follows from the following calculation:
 \[ 1 = \mu_B\lambda_B(d) = tg\lambda_B(d) = t\lambda_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]
\end{proof}
